---
title: "Sinking of the Titanic"
author: "Vincenz Mautner-Markhof"
date: "2020-10-20"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

While exploring the internet for interesting data I came across this dataset which gives information about the passengers on the Titanic, such as the class, age, sex and if they survived or not. In this analysis I will explore the survival rate across different variables.

```{r listing_data, eval=TRUE, warning=FALSE}
titanic <- read_csv("titanic.csv")
glimpse(titanic)
```

```{r}
titanic$Pclass <- as.factor(titanic$Pclass)
titanic$Survived <- as.factor(titanic$Survived)
titanic$Sex <- as.factor(titanic$Sex)
titanic$Embarked <- as.factor(titanic$Embarked)
```

Survival rate

```{r}
prop.table(table(titanic$Survived))
```

Survival rate by gender

```{r}
ggplot(titanic, aes(x = Sex, fill = Survived)) + 
  theme_bw() +
  geom_bar() +
  labs(y = "Passenger Count",
       title = "Titanic Survival Rates by Sex")
```

Survival rate by class of ticket

```{r}
ggplot(titanic, aes(x = Pclass, fill = Survived)) + 
  theme_bw() +
  geom_bar() +
  labs(y = "Passenger Count",
       title = "Titanic Survival Rates by Pclass")
```

Survival rate by class of ticket and gender

```{r}
ggplot(titanic, aes(x = Sex, fill = Survived)) + 
  theme_bw() +
  facet_wrap(~ Pclass) +
  geom_bar() +
  labs(y = "Passenger Count",
       title = "Titanic Survival Rates by Pclass and Sex")
```

Distribution of passenger ages

```{r}
ggplot(titanic, aes(x = Age)) +
  theme_bw() +
  geom_histogram(binwidth = 5) +
  labs(y = "Passenger Count",
       x = "Age (binwidth = 5)",
       title = "Titanic Age Distribtion")
```

Survival rates by age

```{r}
ggplot(titanic, aes(x = Age, fill = Survived)) +
  theme_bw() +
  geom_histogram(binwidth = 5) +
  labs(y = "Passenger Count",
       x = "Age (binwidth = 5)",
       title = "Titanic Survival Rates by Age")
```

```{r}
ggplot(titanic, aes(x = Age, fill = Survived)) +
  theme_bw() +
  facet_wrap(Sex ~ Pclass) +
  geom_density(alpha = 0.5) +
  labs(y = "Age",
       x = "Survived",
       title = "Titanic Survival Rates by Age, Pclass and Sex")
```

Survival rates by age segmented by gender and class of ticket

```{r}
ggplot(titanic, aes(x = Age, fill = Survived)) +
  theme_bw() +
  facet_wrap(Sex ~ Pclass) +
  geom_histogram(binwidth = 5) +
  labs(y = "Age",
       x = "Survived",
       title = "Titanic Survival Rates by Age, Pclass and Sex")
```
